---
title: "msdr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{msdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(msdr)
```
The goal of msdr is to provide conveniance functions to analyse [NOAA Significant Earthquakes dataset](https://www.ngdc.noaa.gov/hazel/view/hazards/earthquake/search).

The package contains three main sets of functions:
1. Cleaning functions
2. Plotting functions (in the form of [ggplot2](https://ggplot2.tidyverse.org/) geoms)
3. Mapping functions

The package contains an example dataset that can be access with msdr::raw_noaa

## Installation and loading

You can install the released version of msdr from [github](https://github.com/ElliottMess/msdr) with:

``` r
devtools::install_github("ElliottMess/msdr")
```

It can then be easily loaded with
``` r
library(msdr)
```


## Cleaning functions

There is two cleaning function:
1. eq_location_clean()
2. eq_clean_data()

eq_location_clean cleans the location names to provide more informative labels.

eq_clean_data formates the whole dataset to implement better data management, including calling eq_location_clean on the location column.


```{r clean}
clean_location_noaa <- eq_location_clean(raw_noaa)

test <- eq_clean_data(raw_noaa)
```


## Plotting functions
Two plotting functions exist:
1. geom_timeline
2. geom_timeline_label

geom_timeline creates a timeline representation of the data by date. 

It plots earthquakes on a timeline ranging from a minimum (xmin) to a maximum (xmax). Each point is an earthquake. The x aesthetic  is a Date class vector of dates of earthquakes.
Size, colour, alpha and y can be provided as optional aesthetic.
y must be a factor and can be used to have one timeline per level of the factor. For instance, each line can be a country.

```{r geom_timeline}
library(ggplot2)
clean <- eq_clean_data(raw_noaa) %>% 
  filter(COUNTRY %in% c("AFGHANISTAN", "NEPAL"))

ggplot(clean, aes(x = DATE, y = COUNTRY, size=EQ_PRIMARY, fill = DEATHS)) +
     geom_timeline(alpha = 0.6, xmin = ymd('2000-01-01'), xmax = ymd('2015-12-31'))
```
geom_timeline_label adds a label to each data point that consists of a line and a text (for instance, the location of the earthquake) attached to each line.
The x aesthetic provides the date vector for the timeline and label provides the labelling vector.
The top_eq convienence function helps find the n_max optional parameter allows to subset the data with the n_max largest earthquakes by magnitude.

```{r geom_timeline_label}
library(dplyr)
library(ggplot2)
library(lubridate)
data <- eq_clean_data(raw_noaa) %>%
        filter(COUNTRY %in% c('USA', 'CHINA'))

labels <- data %>%
  top_eq(n_max = 10, min_date = ymd('2000-01-01'), max_date = ymd('2015-12-31'))

ggplot(data, aes(x = DATE, y = COUNTRY, fill = DEATHS)) +
  geom_timeline(aes(size = EQ_PRIMARY), alpha = 0.25,
                    xmin = ymd('2000-01-01'), xmax = ymd('2015-12-31')) +
  geom_timeline_label(data = labels,
                      mapping = aes(x = DATE, y = COUNTRY, label = LOCATION_NAME))

```
## Mapping functions

The mapping functions allows to explore on a map the datasets. There is two mapping functions:

1. eq_map()
2. eq_create_label()

eq_map() creates maps based on the dataset provided, while eq_create_label() creates more informative label for eq_map.

```{r eq_mapping}

 eq_clean_data(raw_noaa) %>%
     dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
     eq_map(annot_col = "DATE")
```
With labels:
```{r eq_mapping_label}

raw_noaa %>%
  eq_clean_data() %>%
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
  dplyr::mutate(popup_text = eq_create_label(.)) %>%
  eq_map(annot_col = "popup_text")
```

